Flexible Probabilistic Forecasting with Distributional Trees and Forests

### 2018-04-04 pkg/disttree/inst/MIP-Seminar/slides.Rnw
revised slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2123 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 5db251d5602213d07a40d20468176d8d0d8bfafb

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("RainTyrol")
library("gamlss.cens")
gen.cens(NO, type = "left")

@@

\begin{frame}[fragile]
\frametitle{Data modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

@@

\visible<2-6>{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5-6>{
\vspace{-0.8cm}
\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Algorithmic modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3-6>{Tree model\only<3>{\\}\only<4-6>{s}}\only<7->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-1cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain now, humdity\\
Y = rain 24h}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.2, 3.5) {};
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 20pt, line width = 3pt] (subset3) at (1.6, 3.2) {};
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
\end{tikzpicture}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}

@@

%\vspace{-0.29cm}

@@

par(mar=c(3,3,2,5))

@@

lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)
mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)

@@

%\hspace{-1cm}

@@

@@

lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)

@@

mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)

@@

@@

mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)

@@

lines(x = x, y = ytree_up, col = pal["forest"], lwd=7, lty = 2)

@@

mtext("rain", side = 4, las = TRUE, at = 15, line = 0.5, cex = 3, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 8, line = 0.5, cex = 3, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.4, cex = 3)

@@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{minipage}{0.78\textwidth}
\begin{center}
\vspace{0.2cm}
\visible<2->{
Model: \code{disttree(y ~ x)}\\
}
\vspace{-0.2cm}
\only<2-3>{
%\setkeys{Gin}{width=0.\linewidth, height=0.68\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE, width = 7>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

@@

\center
\visible<5->{
Algorithmic \\
Modelling\\
}

@@

\visible<6->{
Data \\
Modelling
}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Application for one stations:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for all 95 station:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
\end{frame}

@@

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(dem, col = rev(gray.colors(100)),
                main="Stations in Tyrol", cex.main=1.7,
                ylab = "Latitude", xlab = "Longitude",
                cex.lab = 1.4,
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(tirol, add = TRUE)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.25, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.4,-1.2))

>>>>

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("RainTyrol")
library("gamlss.cens")
gen.cens(NO, type = "left")

@@

\begin{frame}[fragile]
\frametitle{Data modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$.\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

@@

\visible<2-6>{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5-6>{
\vspace{-1.3cm}

\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Algorithmic modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3-6>{Tree model\only<3>{\\}\only<4-6>{s}}\only<7->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-0.96cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain today, humidity\\
Y = rain tomorrow}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
}
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\vspace{0.1cm}

@@

%\vspace{0.01cm}

@@

par(mar=c(3,3,2,5.5))

@@

lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain tomorrow", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)

@@

@@

%\vspace{0.01cm}
\hspace{-0.2cm}

@@

lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)

@@

lines(x = x, y = ytree_up, col = pal["tree"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain tomorrow", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)

@@

%\vspace{0.01cm}
\hspace{-0.3cm}

@@

lines(x = x, y = ytree_up, col = "blue", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain tomorrow", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)

@@

lines(x = x, y = ytree_up, col = "orange", lwd=7)

@@

mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("today", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain tomorrow", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)

@@

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{minipage}{0.78\textwidth}
\begin{center}
\vspace{0.2cm}
\visible<2->{
Model: \code{disttree(y $\sim$ x)}\\
}
\vspace{-0.2cm}
\only<2-3>{
%\setkeys{Gin}{width=0.\linewidth, height=0.68\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE, width = 7>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

@@

\center
\visible<5->{
Algorithmic \\
Modeling\\
}

@@

\visible<6->{
Data \\
Modeling
}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for all 95 stations:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item Compare to commonly used heteroscedastic censored Gaussian models.
\end{itemize}
\end{frame}

@@

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(dem, col = rev(gray.colors(100)),
                main="Stations in Tyrol", cex.main=1.7,
                ylab = "Latitude", xlab = "Longitude",
                cex.lab = 1.3,
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(tirol, add = TRUE)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.25, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.4,-1.2))
